---
title: "2019-03-23_Harris_29"
author: "Nicolas May"
date: "4/3/2019"
output: 
  html_document:
    theme: sandstone
    highlight: zenburn
    toc: true
    toc_float: true
---

# BUILD DATASET

## Load Requirements

### Load Packages

```{r}
suppressMessages( library(dplyr)   )
suppressMessages( library(stringr) )
```

### Load Useful Globals / Functions

```{r}
source("~/Desktop/config.R")
source("~/Desktop/helpers.R")

GET_API_DATA <- TRUE
```


## Get Data

### Define Data Fields/Forms

UDS 3

```{r}
# Header form
fields_u3_hd_vct <-
  c(
    "ptid"
    , "form_date"
  )
# A1 form
fields_u3_a1_vct <-
  c(
    "sex"
    , "hispanic"
    , "race"
    , "residenc"
  ) %>% 
  c(., paste0("fu_", .), paste0("tele_", .))
# D1 form
fields_u3_d1_vct <-
  c(
    # Normal cognition
    "normcog"
    # Dementia
    , "demented"
    # Syndromes
    , "amndem"
    # , "pca"
    # , "ppasyn"
    # , "ftdsyn"
    # , "lbdsyn"
    # , "namndem"
    # Etiologies
    , "alzdis"
    , "alzdisif"
    , "cvd"
    , "cvdif"
  ) %>% 
  c(., paste0("fu_", .), paste0("tele_", .))

# Collapse `fields_u3_*_vct`s
fields_u3_vct <-
  c(
    fields_u3_hd_vct
    , fields_u3_a1_vct
    , fields_u3_d1_vct
  )
fields_u3 <- fields_u3_vct %>% paste(collapse = ",")
```

MiNDSet Registry

```{r}
# Registration Form
fields_ms_reg_vct <-
  c(
    "subject_id"
  )
# Demographics Form
fields_ms_dem_vct <-
  c(
    "race_value"
    , "sex_value"
  )
# Medical Visit Form
fields_ms_med_vct <-
  c(
    "exam_date"
  )

fields_ms_vct <-
  c(
    fields_ms_reg_vct
    , fields_ms_dem_vct
    , fields_ms_med_vct
  )
fields_ms <- fields_ms_vct %>% paste(collapse = ",")
```


### Retrieve Data via REDCap API

UDS 3

```{r}
if (GET_API_DATA) {
  json_u3 <- rc_api_get(uri    = REDCAP_API_URI,
                        token  = REDCAP_API_TOKEN_UDS3n,
                        fields = fields_u3,
                        raw    = TRUE)
}
df_u3 <- jsonlite::fromJSON(json_u3) %>% na_if("")
```

MiNDSet Registry

```{r}
if (GET_API_DATA) {
  json_ms <- rc_api_get(uri    = REDCAP_API_URI,
                        token  = REDCAP_API_TOKEN_MINDSET,
                        fields = fields_ms,
                        raw    = TRUE)
}
df_ms <- jsonlite::fromJSON(json_ms) %>% na_if("")
```


## Process Data

### Clean Data

UDS 3

```{r}
df_u3_cln <- df_u3 %>% 
  # remove useless `redcap_event_name` field
  select(-redcap_event_name) %>% 
  # clean out records missing `form_date`s
  filter(!is.na(form_date)) %>% 
  # clean out DDE records
  filter(str_detect(ptid, pattern = "^UM\\d{8}$"))
```

MiNDSet Registry

```{r}
df_ms_cln <- df_ms %>% 
  # remove useless `redcap_event_name` field
  select(-redcap_event_name) %>% 
  # clean out records missing `exam_date`s
  filter(!is.na(exam_date)) %>% 
  # clean out non-UMMAP IDs
  filter(str_detect(subject_id, pattern = "^UM\\d{8}$"))
```


### Mutate Data

UDS 3

```{r}
df_u3_cln_mut <- df_u3_cln %>% 
  # coalesce IVP / FVP / TVP fields
  coalesce_ift_cols() %>% 
  # retype fields appropriately
  mutate(form_date = as.Date(form_date)) %>% 
  # get most recent visit only
  get_visit_n(ptid, form_date, Inf) %>% 
  # simplify diagnosis
  mutate(uds_dx = case_when(
    normcog    == 0 &
      demented == 1 &
      amndem   == 1 &
      alzdis   == 1 &
      alzdisif == 1 ~ "Alzheimer's Disease",
    normcog    == 0 &
      demented == 1 &
      cvd      == 1 &
      cvdif    == 1 ~ "Vascular Dementia",
    TRUE ~ NA_character_
  ))
### Alzheimer's Disease:
###   normcog  == 0 &
###   demented == 1 &
###   amndem   == 1 &
###   alzdis   == 1 &
###   alzdisif == 1
### Vascular Dementia:
###   normcog  == 0 &
###   demented == 1 &
###   cvd      == 1 &
###   cvdif    == 1
```

MiNDSet Registry

```{r}
df_ms_cln_mut <- df_ms_cln %>% 
  # retype fields appropriately
  mutate(exam_date = as.Date(exam_date)) %>% 
  # change `race_value` to match UDS 3 `race`
  mutate(race_ms = case_when(
    race_value == 1 ~ 1L,
    race_value == 2 ~ 2L,
    race_value == 3 ~ 5L,
    race_value == 5 ~ 50L,
    race_value == 6 ~ 99L,
    TRUE ~ NA_integer_
  )) %>% 
  mutate(sex_ms = sex_value) %>% 
  # deselect unneeded fileds
  select(-race_value, -sex_value)
```


### Filter Data

UDS 3

```{r}
df_u3_cln_mut_flt <- df_u3_cln_mut %>% 
  filter(!is.na(uds_dx))
```

### Join Data

UDS 3 + MiNDSet Registry

```{r}
df_u3_ms <- left_join(x = df_u3_cln_mut_flt, 
                      y = df_ms_cln_mut,
                      by = c("ptid" = "subject_id",
                             "form_date" = "exam_date"))
# Coalesce sex and race fields
df_u3_ms_mut <- df_u3_ms %>% 
  mutate(sex = coalesce(as.integer(sex), as.integer(sex_ms))) %>% 
  mutate(race = coalesce(as.integer(race), as.integer(race_ms))) %>% 
  select(-sex_ms, -race_ms)

# Clean up 
df_u3_ms_mut_cln <- df_u3_ms_mut %>% 
  mutate(hispanic = case_when(
    hispanic == 1 ~ 1L,
    TRUE ~ 0L
  )) %>% 
  select(-normcog
         , -demented
         , -amndem
         , -alzdis
         , -alzdisif
         , -cvd
         , -cvdif)

# Hash pt. IDs
df_u3_ms_mut_cln_hsh <- df_u3_ms_mut_cln %>% 
  mutate(ptid = openssl::md5(ptid))
```

Convert raw data to labelled data

```{r}
df_u3_ms_mut_cln_hsh_lbl <- df_u3_ms_mut_cln_hsh %>% 
  # sex
  mutate(sex = case_when(
    sex == 1 ~ "Male",
    sex == 2 ~ "Female",
    TRUE ~ NA_character_
  )) %>% 
  # race
  mutate(race = case_when(
    race == 1 ~ "White",
    race == 2 ~ "Black or African American",
    race == 3 ~ "American Indian or Alaska Native",
    race == 4 ~ "Native Hawaiian or other Pacific Islander",
    race == 5 ~ "Asian",
    race == 50 ~ "Other",
    race == 99 ~ "Unknown",
    TRUE ~ NA_character_
  )) %>% 
  # hispanic ethnicity
  mutate(hispanic = case_when(
    hispanic == 1 ~ "Hispanic ethnicity",
    TRUE ~ NA_character_
  )) %>% 
  # residence
  mutate(residenc = case_when(
    residenc == 1 ~ "Single- or multi-family private residence",
    residenc == 2 ~ "Retirement community or independent group living",
    residenc == 3 ~ "Assisted living, adult family home, or boarding home",
    residenc == 4 ~ 
      "Skilled nusring facility, nursing home, hospital, or hospice",
    residenc == 9 ~ "Unknown",
    TRUE ~ NA_character_
  )) 
```



## Write Data

### Write as CSV

```{r}
# raw data
readr::write_csv(df_u3_ms_mut_cln_hsh,
                 "df_u3_ms_mut_cln_hsh.csv",
                 na = "")
# label data
readr::write_csv(df_u3_ms_mut_cln_hsh_lbl, 
                 "df_u3_ms_mut_cln_hsh_lbl.csv", 
                 na = "")
```


## Summarize Data

```{r}
(uds_dx_summ <- df_u3_ms_mut_cln_hsh_lbl %>% 
   group_by(sex, race, hispanic, residenc, uds_dx) %>% 
   tally() %>% 
   arrange(uds_dx, sex, race, hispanic, residenc) %>% 
   select(uds_dx, sex, race, hispanic, residenc, everything()))
```
















